Pool Model Paper

### 2023-12-05 pool_model.tex
caption size formatting

In the described simplified version of the real-life system three possible bacteria states are mentioned corresponding to which process is happeng: adaplation, growth or death.

@@

    \caption{{\footnotesize Schematic representation of exchange between three states or pools.}}

@@

    \caption{{
        \footnotesize The time dependence of the total bacterial count $N(t) = \sum_i \sum_k \mbx_i^k = L(t) + G(t) + D(t)$ is obtained as a sum of bacteria in all states.

@@

    }}

@@

    \caption{{
        \footnotesize (a) The time dependence of the total bacterial count $N$ for the model including the back-flow to the lag phase due to temperature stres (\ref{eq:ODE_tempshift}).

@@

\begin{figure}[H]
    \begin{center}
    \includegraphics[width=0.67\textwidth]{Figures/pool_model_2pools_resource_competition.pdf}
    \caption{{
        \footnotesize The competition of the two bacterial species due to the common limiting resource calculated using Eq. (\ref{eq:model_2sp_resource_comp}).
        The red curve describes the growth of the bacteria $A$ $N_A = \sum_{k} x_A^k = L_A+G_A$ with model parameters $\lambda_A=10^{-2}$, $\alpha_A=3.0$.
        The blue curve describes the growth of the bacteria $B$ $N_B = \sum_{k} x_B^k = L_B+G_B$ with model parameters $\lambda_A=5\cdot 10^{-2}$, $\alpha_A=2.5$.
        The purple line shows the total bacterial count of the system  $N(t)=\sum_i\sum_k\mbx_i^k=N_A+N_B$.
        The gray line denotes the dynamic of the resource pool scaled with its capacity $N_t=10^4$.
    }}
    \label{fig:2pool_resource_2sp}
    \end{center}
\end{figure}
As one can see in Figure~\ref{fig:2pool_resource_2sp} even without direct interaction between species the bacterial growth was inhibited due to the limitation of the resource.
Here species $A$ was able to supress species $B$ due to the higher values of the growth constant $\alpha$.
%
\cmt{But LV for several bacteria states in literature? Didn't find!}
%
%
%
% ------------------------------------------------------------------------------------------------
\subsection{Interspecies Competition}
However, except the inhibition due to resource competition, bacteria developed different others active competition mechanisms.
One of the examples of such mechanisms is a production of the antimicrobial compounds, inhibitors or toxins~\cite{wloch-salamon_effect_2008, chao_structured_1981}.
%
%
% ................................................................................................
\subsubsection{Toxins Production}
Now extend our model where bacteria not only consume nutritions from resource pool $R$ but one of the species also actively produce the inhibiting compounds $T$. 
It is the accumulation of the bacteria waste which is responsible for the spoilage of food products. 
In order to take this into account we add an equation for the waste production:
\begin{eqnarray}
    \dot{L} &=& - \lambda_f R L + \mathcal{B}(R)G\\
    \dot{G} &=& \lambda_f R L +\alpha R G-\mathcal{B}(R)G\\
    \dot{R} &=&-\frac{\alpha}{N_t} R G\\
    \dot{T} &=& \mathcal{F}(R,G).
\end{eqnarray}
The function $\mathcal{F}$ denotes the model for the antibiotic production rate. 
We shall discuss two possible models:
\begin{eqnarray}
    \mathcal{F}_1(R,G)&=&\kappa G\\
    \mathcal{F}_2(R,G)&=&\kappa\alpha R G=-\kappa N_t\dot{R}
\end{eqnarray}
We would like to test the following model:
\begin{equation}
    \begin{cases}
        \dot{L}_A = - \lambda_A R L_A\\
        \dot{L}_B = - \lambda_B R L_B \\
        \dot{G}_A = \lambda_A R L_A + \alpha_A R G_A - \frac{\mu}{\kappa N_t} T G_A\\
        \dot{G}_B = \lambda_B R L_B + \alpha_B R G_B\\
        \dot{R} = -\frac{\alpha_A}{N_t} R G_A-\frac{\alpha_B}{N_t} R G_B\\
        \dot{T} = \mathcal{F}_{1,2} (R, G_B) \\
    \end{cases}
    \label{eq:model_2sp_toxin}
\end{equation}

@@

    \caption{{
        \footnotesize The two species system competing for the common nutritient $R$ with production of the toxin $T$ (brown line) calculated using Eq. (\ref{eq:model_2sp_toxin}).

@@

    \caption{{
        \footnotesize The two species system where species inhibit each other defined with equations~(\ref{eq:2sp_1pool_inhibit_unitless1}-\ref{eq:2sp_1pool_inhibit_unitless2}).

@@

    \caption{{\footnotesize The list of the parameters of the system.}}

@@

    \caption{{\footnotesize The list of the initial values of the system.}}

@@

    \caption{{
        \footnotesize The output of the model (\ref{eq:spatial_limit_F}, \ref{eq:spatial_limit_H}) for given parameters and initial valued presented in Tables \ref{tab:spatial_limit_param}, \ref{tab:spatial_limit_init}, respectively.

>>>>

In the described simplified version of the real-life system three possible bacteria states are mentioned corresponding to which process is happening: adaptation, growth or death.

@@

    \caption{Schematic representation of exchange between three states or pools.}

@@

    \caption{
        The time dependence of the total bacterial count $N(t) = \sum_i \sum_k \mbx_i^k = L(t) + G(t) + D(t)$ is obtained as a sum of bacteria in all states.

@@

    }

@@

    \caption{
        (a) The time dependence of the total bacterial count $N$ for the model including the back-flow to the lag phase due to temperature stres (\ref{eq:ODE_tempshift}).

@@

\begin{figure}[H]
    \begin{center}
    \includegraphics[width=0.67\textwidth]{Figures/pool_model_2pools_resource_competition.pdf}
    \caption{
        The competition of the two bacterial species due to the common limiting resource calculated using Eq. (\ref{eq:model_2sp_resource_comp}).
        The red curve describes the growth of the bacteria $A$ $N_A = \sum_{k} x_A^k = L_A+G_A$ with model parameters $\lambda_A=10^{-2}$, $\alpha_A=3.0$.
        The blue curve describes the growth of the bacteria $B$ $N_B = \sum_{k} x_B^k = L_B+G_B$ with model parameters $\lambda_A=5\cdot 10^{-2}$, $\alpha_A=2.5$.
        The purple line shows the total bacterial count of the system  $N(t)=\sum_i\sum_k\mbx_i^k=N_A+N_B$.
        The gray line denotes the dynamic of the resource pool scaled with its capacity $N_t=10^4$.
    }
    \label{fig:2pool_resource_2sp}
    \end{center}
\end{figure}
As one can see in Figure~\ref{fig:2pool_resource_2sp} even without direct interaction between species the bacterial growth was inhibited due to the limitation of the resource.
Here species $A$ was able to supress species $B$ due to the higher values of the growth constant $\alpha$.
%
\cmt{But LV for several bacteria states in literature? Didn't find!}
%
%
%
% ------------------------------------------------------------------------------------------------
\subsection{Interspecies Competition}
However, except the inhibition due to resource competition, bacteria developed different others active competition mechanisms.
One of the examples of such mechanisms is a production of the antimicrobial compounds, inhibitors or toxins~\cite{wloch-salamon_effect_2008, chao_structured_1981}.
%
%
% ................................................................................................
\subsubsection{Toxins Production}
Now extend our model where bacteria not only consume nutritions from resource pool $R$ but one of the species also actively produce the inhibiting compounds $T$. 
It is the accumulation of the bacteria waste which is responsible for the spoilage of food products. 
In order to take this into account we add an equation for the waste production:
\begin{eqnarray}
    \dot{L} &=& - \lambda_f R L + \mathcal{B}(R)G\\
    \dot{G} &=& \lambda_f R L +\alpha R G-\mathcal{B}(R)G\\
    \dot{R} &=&-\frac{\alpha}{N_t} R G\\
    \dot{T} &=& \mathcal{F}(R,G).
\end{eqnarray}
The function $\mathcal{F}$ denotes the model for the antibiotic production rate. 
We shall discuss two possible models:
\begin{eqnarray}
    \mathcal{F}_1(R,G)&=&\kappa G\\
    \mathcal{F}_2(R,G)&=&\kappa\alpha R G=-\kappa N_t\dot{R}
\end{eqnarray}
We would like to test the following model:
\begin{equation}
    \begin{cases}
        \dot{L}_A = - \lambda_A R L_A\\
        \dot{L}_B = - \lambda_B R L_B \\
        \dot{G}_A = \lambda_A R L_A + \alpha_A R G_A - \frac{\mu}{\kappa N_t} T G_A\\
        \dot{G}_B = \lambda_B R L_B + \alpha_B R G_B\\
        \dot{R} = -\frac{\alpha_A}{N_t} R G_A-\frac{\alpha_B}{N_t} R G_B\\
        \dot{T} = \mathcal{F}_{1,2} (R, G_B) \\
    \end{cases}
    \label{eq:model_2sp_toxin}
\end{equation}

@@

    \caption{
        The two species system competing for the common nutritient $R$ with production of the toxin $T$ (brown line) calculated using Eq. (\ref{eq:model_2sp_toxin}).

@@

    \caption{
        The two species system where species inhibit each other defined with equations~(\ref{eq:2sp_1pool_inhibit_unitless1}-\ref{eq:2sp_1pool_inhibit_unitless2}).

@@

    \caption{The list of the parameters of the system.}

@@

    \caption{The list of the initial values of the system.}

@@

    \caption{
        The output of the model (\ref{eq:spatial_limit_F}, \ref{eq:spatial_limit_H}) for given parameters and initial valued presented in Tables \ref{tab:spatial_limit_param}, \ref{tab:spatial_limit_init}, respectively.
